---
title: "Population Ecology HW1"
author: "Tess Hooper"
date: "1/15/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

# Question 1. Why might a plant or animal be patchily distributed? 
1. Availability of resources (e.g. water, food) might affect where plants or animals are distributed. In this case I would expect the location of high-density patches to be inconsistent from year to year, as the resources may change. For instance, if certain foraging plants are abundant in one area, then that is where we might find ungulates in a given year. However, if they deplete that resource then that group of ungulates may need to move locations. 
2. 
3. 

```{r load packages, echo = FALSE, message = FALSE, warning=FALSE}

library(tidyverse)

```

```{r read in data, echo = FALSE, message = FALSE}

grizzly <- read_csv("grizzlydata.csv")
swallenia2 <- read_csv("Swallenia (1).csv")

```

<br>

# Question 2: Eureaka Dune Grass
```{r, A: two-sample t-test, echo = FALSE, message = FALSE}

# A. Change in mean abundance from 2009 to 2010 using a two-sample t-test


swallenia_ttest <- t.test(swallenia2$count_2009, swallenia2$count_2010)

swallenia_ttest

# Null Hypothesis: There is no significant change in means between 2009 - 2010
# Altnerative Hypothesis: There is a significance change in means between 2009 - 2010

## If p-value is greater than 0.05 (alpha) then we accept the null. If it is less than 0.05 we reject the null. 


#### Results ####

#data:  swallenia2$count_2009 and swallenia2$count_2010
#t = -0.81791, df = 17.37, p-value = 0.4245
#alternative hypothesis: true difference in means is not equal to 0
#95 percent confidence interval:
# -141.06385   62.15476
#sample estimates:
#mean of x mean of y 
# 61.90909 101.36364 

#### Answer ####

## For this analysis I chose 0.05 as my alpha value because this is a common standard that we have used in the past (e.g. Allison's class). The t-test resulted in a p-value of 0.4245. Based on this analysis, there was no significant change in mean abundance of Swallenia between 2009 and 2010, since this p-value was greater than 0.05. 

```
For this analysis I chose 0.05 as my alpha value because this is a common standard that we have used in the past (e.g. Allison's class). The t-test resulted in a p-value of 0.4245. Based on this analysis, there was no significant change in mean abundance of Swallenia between 2009 and 2010, since this p-value was greater than 0.05.

<br>

```{r, B: two-sample paired t-test, echo = FALSE, message = FALSE }

# B. Change in mean abundance using a paired two sample t-test




```

C. Which analysis is more appropriate? 

<br>

D. Paragraph to the park superintendant describing my findings

<br>
<br>

# Question 3: Yellowstone Grizzly Bears
```{r, A: bear population, echo = FALSE, message = FALSE}

# A. Did the grizzly population decline between 1959 - 1968? Use a linear regression? 




```

<br>

```{r, B: rate of decline, echo = FALSE, message=FALSE}

# B. Did the population continue to decline from 1969 to 1978? Was the decline faster than the period prior to the dump closures? 



```

<br>

```{r, C: magnitude of populatio trends, echo=FALSE, message=FALSE}

# C. Did the population size continue to change after 1978? What was the direction and magnitude of the trend? 



```


